---
title: "Massachusetts Table 2"
author: "Julie Ghekas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

I am trying to replicate Table 2 in the Sommers, et al. paper.  According to the paper,
within Massachusetts, the age distribution is 33.2% 20-34 yr olds, 26.3% 35-44, 24.0% 
45-54, and 16.5% 55-64 yr olds.  Here, I try looking at 2001-2005, 2001-2006, 2005, 
and 2006 as the baseline years.  I also try unweighted and weighted means for 
calculating the overall percentage for Massachusetts.  The full distribution is listed
for the unweighted means, while the weighted means are reported only for 20-34 year 
olds. 

```{r}
stopifnot(file.exists(prepped_data_file <- paste0(datadir, "created/mass_prep.RData")))
load(prepped_data_file)
```

Using just 2005 data:

```{r}
pop05 = subset(popall, popall$year==2005)
mat1 = aggregate(cbind(a2034, a3544, a4554, a5564)~county + state, sum, data=pop05)
ages = cbind(mat1[,1:2], 100*t(apply(mat1[,3:6], 1, prop.table)))
mas1 = subset(ages, ages$state==25)
#unweighted mean
lapply(mas1, mean)
#weighted mean
mat1 = subset(mat1, mat1$state==25)
sum(mat1[,3])/sum(mat1[,c(3:6)])
```

For 2001-2005

```{r}
pop05 = subset(popall, popall$year>=2001 & popall$year<=2005)
mat1 = aggregate(cbind(a2034, a3544, a4554, a5564)~county + state, sum, data=pop05)
ages = cbind(mat1[,1:2],100*t(apply(mat1[,3:6],1,prop.table)))
mas1 = subset(ages, ages$state==25)
#unweighted mean
lapply(mas1,mean)
#weighted mean for 20-34 year olds
mat1 = subset(mat1, mat1$state==25)
sum(mat1[,3])/sum(mat1[,c(3:6)])
```

For 2006:

```{r}
pop06 = subset(popall, popall$year==2006)
mat1 = aggregate(cbind(a2034,a3544,a4554,a5564)~county+state,sum,data=pop06)
ages = cbind(mat1[,1:2],100*t(apply(mat1[,3:6],1,prop.table)))
mas1 = subset(ages, ages$state==25)
#unweighted mean
lapply(mas1,mean)
#weighted mean
mat1 = subset(mat1, mat1$state==25)
apply(mat1, 2, sum)/sum(mat1[,3:6])
```

For 2001-2006:

```{r}
pop06= subset(popall, popall$year>=2001 & popall$year<=2006)
mat1 = aggregate(cbind(a2034, a3544, a4554, a5564)~county + state, sum, data=pop06)
ages = cbind(mat1[,1:2],100*t(apply(mat1[,3:6],1,prop.table)))
mas1 = subset(ages, ages$state==25)
#unweighted mean
lapply(mas1,mean)
#weighted mean for Mass
matmas = subset(mat1, mat1$state==25)
apply(matmas, 2, sum)/sum(matmas[,3:6])
#weighted mean for US (including Mass)
apply(mat1, 2, sum)/sum(mat1[,3:6])
#weighted mean for US (w/o Mass)
matno = subset(mat1, mat1$state!=25)
apply(matno, 2, sum)/sum(matno[,3:6])


racesexmat = aggregate(cbind(count=a2034+a3544+a4554+a5564)~racesex+county+state,
                       sum, data=pop06)
racesmat = racesexmat
racesmat$race=recode(racesexmat$racesex,"1:2=1; 3:4=2; 5:8=3")
racesmat$race[racesmat$race==1]="white"
racesmat$race[racesmat$race==2]="black"
racesmat$race[racesmat$race==3]="otherrace"
races=aggregate(count~race+county+state,sum,data=racesmat)
races=dcast(data=races,state+county~race, value.var='count')
#mass weighted race
race = subset(races, races$state==25)
apply(race, 2, sum)/sum(race[,3:5])
#no mass weighted race
raceno = subset(races, races$state!=25)
apply(raceno, 2, sum)/sum(raceno[,3:5])
#all weighted race
apply(races, 2, sum)/sum(races[,3:5])

sexmat=racesexmat
sexmat$sex=recode(racesexmat$racesex, "1=1; 2=2; 3=1; 4=2; 5=1; 6=2; 7=1; 8=2")
sexmat$sex[sexmat$sex==1]="male"
sexmat$sex[sexmat$sex==2]="female"
sexes=aggregate(count~sex+county+state,sum,data=sexmat) 
sexes=dcast(data=sexes,state+county~sex, value.var = 'count')
#mass
sexmass = subset(sexes, sexes$state==25)
apply(sexmass, 2, sum)/sum(sexmass[,3:4])
#all us
apply(sexes, 2, sum)/sum(sexes[,3:4])
#all us less mass
sexno = subset(sexes, sexes$state!=25)
apply(sexno, 2, sum)/sum(sexno[,3:4])

hispmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~hisp+county+state,sum,data=pop06)
hispmat$hisp=recode(hispmat$hisp, '2=1; 1=2; 9=2')
hispmat$hisp[hispmat$hisp==1]="hisp"
hispmat$hisp[hispmat$hisp==2]='nothisp'
hisps=aggregate(count~hisp+county+state,sum,data=hispmat) 
hisps=dcast(data=hisps,state+county~hisp, value.var='count')
#mass
hispmass = subset(hisps, hisps$state==25)
apply(hispmass, 2, sum)/sum(hispmass[,3:4])
#us
apply(hisps, 2, sum)/sum(hisps[,3:4])
#us less mass
hispno = subset(hisps, hisps$state!=25)
apply(hispno, 2, sum)/sum(hispno[,3:4])


```

Above matches what is reported in the table (2001-2006, weighted means).
The rounding on the race category seems a little off (but may be a result of adding 
to 100).  The Latino category seems a little off, but less than 1%.
Rest of US is US without Mass (which makes sense).  Again, excepting some small 
rounding differences, I didn't notice many large differences.  The Latino category 
again does not calculate correctly.

Trying for the arf variables:

```{r}
arfcalc = melt(arfall, id=c('state', 'county'))
arfcalc$var = substr(arfcalc$variable, 1, 3)
arfcalc$year = as.numeric(substr(arfcalc$variable, 4, 5))
arfcalc$year=recode(arfcalc$year, "7=2007;0=2000; 1=2001; 2=2002; 3=2003; 4=2004; 
                    5=2005; 6=2006; 8=2008; 9=2009; 10=2010; 11=2011")
arfcalc = dcast(data=arfcalc, state+county+year~var)
arf06 = subset(arfcalc, arfcalc$year>=2001 & arfcalc$year<=2006)
popbyyr = aggregate(cbind(popn=a2034+a3544+a4554+a5564)~county+state+year, sum, 
                    data=pop06)
arfpop = merge(arf06, popbyyr, by=c('state','county','year'))
#need to go back and relook at this.... double check missing ARF variables
#double check that the difference in the 3230 ARF counties and 3139/3141 population
#counties

#unweighted mean
#mass
arfpopmass = subset(arfpop, arfpop$state==25)
apply(arfpopmass, 2, function(x) mean(x, na.rm=T))
#us
apply(arfpop, 2, function(x) mean(x, na.rm=T))
#us less mass
arfpopno = subset(arfpop, arfpop$state!=25)
apply(arfpopno, 2, function(x) mean(x, na.rm=T))

#weighted mean
#us
apply(arfpop[c(4, 5, 7)], 2, function(x) sum(x*arfpop$popn)/sum(arfpop$popn))
arf05 = subset(arfpop, arfpop$year>=2005 & arfpop$year<=2006)
sum(arf05$ins * arf05$popn)/sum(arf05$popn)
#mass
apply(arfpopmass[c(4, 5, 7)], 2, function(x) 
  sum(x*arfpopmass$popn)/sum(arfpopmass$popn))
arf05mass = subset(arf05, arf05$state==25)
sum(arf05mass$ins * arf05mass$popn) / sum(arf05mass$popn)
#us less mass
apply(arfpopno[c(4, 5, 7)], 2, function(x) sum(x*arfpopno$popn)/sum(arfpopno$popn))
arf05no = subset(arf05, arf05$state!=25)
sum(arf05no$ins * arf05no$popn) / sum(arf05no$popn)


```

From this, I couldn't get the insurance variables to match up.  I also was not able to 
identify the 2001-2004 versions of those variables, so that may contribute to the lack 
of match up.  Interestingly, the insurance values for the years I could find had an 
uninsurance rate lower in Massachusetts than reported (12.3%) and an unsuriance rate 
high in the rest of the US (20.1%).  

From the description in the Sommers paper, it looks like the % Hispanic may have come 
from the ARF and not from the CDC's CMF.  This is something to look into.

Now, to check for the control group values given in Table 2:

```{r}

```