---
title: "Data prep for Mass. PS modeling script"
author: "Julie Ghekas, Ben Hansen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

To begin, I read in my data from the CDC.

```{r}
if (is.null(datadir)) #reproduce results from when below were hard-coded; see dot-Renviron-template.txt
{
datadir = "~/Downloads/" 
mortdatadir = "~/Downloads/mort/"
buffsize = 2e5
}
if (!file.exists(paste0(datadir, "created"))) stop(paste0("Have to create 'created' subdir of ", datadir))

mort99 = read.fwf(paste0(mortdatadir, 'MORT9913.txt'), width=c(2,3,4,1,1,2,4,3,4), buffersize=buffsize)
colnames(mort99) <- c('state','county','year','racesex','hisp','age','ICD10cause','cause','deaths')
pop99 = read.fwf(paste0(datadir, 'pop9913.txt'), width=c(2,3,4,1,1,rep(8,14),25,1), buffersize=buffsize)
colnames(pop99) = c('state','county','year','racesex','hisp','birth','l1','a14','a59','a1014','a1519','a2024','a2534','a3544','a4554','a5564','a6574','a7584','a85','name','type')
arf = read.csv(paste0(datadir, 'arfarfsixten.csv'), header=T)
colnames(arf) = c('state','county','med610','medinc10','medinc06','pov10','pov06','ins10','ins06','unemp10','unemp06')
arf08 = read.csv(paste0('./', 'arf08vars.csv'), header=T)
arf12 = read.csv(paste0('./', 'arf12vars.csv'), header=T) 
colnames(arf08) = c('state','county','inc07','inc06','inc05','inc04','inc03','inc02',
                    'inc01','inc00',
                    'pov07','pov06','pov05','pov04','pov03','pov02','pov01','pov00',
                    'ins05','emp07','emp06','emp05','emp04','emp03','emp02','emp01','emp00')
colnames(arf12) = c('state','county','inc11','inc10','inc09','inc08','inc07','inc06','inc05',
                    'pov11','pov10','pov09','pov08','pov07','pov06','pov05','ins10','ins09',
                    'ins08',
                    'ins07','ins06','emp11','emp10','emp09','emp08','mep07','emp06','emp05')
arf12 = arf12[,-c(7,8,9,14,15,16,26,27,28)]
arfall = merge(arf08, arf12, by=c('state','county'), all=T)
```

I then subset the data to include only the years I want for the Massachusetts replication, only counties, and add in a variable for an age group (simply the addition of two age groups).

```{r}
pop06 = subset(pop99, pop99$year>=2001 & pop99$year<=2006)
pop06 = subset(pop06, pop06$type==3)
pop06$a2034 = pop06$a2024 + pop06$a2534

mort06 = subset(mort99, mort99$year>=2001 & mort99$year<=2006)
mort06 = subset(mort06, mort06$age>=09 & mort06$age<=13)

```

The following section was added to provide a mortality rate for all years, to help with the analysis of the comparison post-treatment.

```{r}
popall = subset(pop99, pop99$type==3)
popall$a2034 = popall$a2024 + popall$a2534
mortall = subset(mort99, mort99$age>=09 & mort99$age<=13)

```

As that was the last use of `pop99` in this script:
```{r}
rm(pop99, mort99, arf08, arf12)
```

Save the results of this script, along with a record of the state of
this script on the local file server at the time that the save was
made.

```{r}
mass_prep_git_state <- list()
mass_prep_git_state$latest_commit <- system("git log -n 1 --pretty=format:%h -- mass_prep.Rmd",
                                  intern=TRUE)
mass_prep_git_state$diff_not_staged <- system("git diff -- mass_prep.Rmd",
                                             intern=TRUE)
mass_prep_git_state$diff_staged <- system("git diff --staged mass_prep.Rmd",
                                             intern=TRUE)
save(popall,mortall,pop06,mort06,arf,arfall,mass_prep_git_state,
     file=paste0(datadir, "created/mass_prep.RData"))
```
 
