---
title: "Data prep for Mass. PS modeling script"
author: "Julie Ghekas, Ben Hansen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

To begin, I read in my data from the CDC.

```{r}
if (is.null(datadir)) #reproduce results from when below were hard-coded; see dot-Renviron-template.txt
{
datadir = "~/Downloads/" 
mortdatadir = "~/Downloads/mort/"
buffsize = 2e5
}
if (!file.exists(paste0(datadir, "created"))) stop(paste0("Have to create 'created' subdir of ", datadir))

mort99=read.fwf(paste0(mortdatadir, 'MORT9913.txt'),width=c(2,3,4,1,1,2,4,3,4), buffersize=buffsize)
colnames(mort99)=c('state','county','year','racesex','hisp','age','ICD10cause','cause','deaths')
pop99=read.fwf(paste0(datadir, 'pop9913.txt'),width=c(2,3,4,1,1,rep(8,14),25,1), buffersize=buffsize)
colnames(pop99)=c('state','county','year','racesex','hisp','birth','l1','a14','a59','a1014','a1519','a2024','a2534','a3544','a4554','a5564','a6574','a7584','a85','name','type')
arf<-read.csv(paste0(datadir, 'arfarfsixten.csv'),header=T)
colnames(arf)<-c('state','county','med610','medinc10','medinc06','pov10','pov06','ins10','ins06','unemp10','unemp06')
library(reshape2)
library(car)
library(RItools)
```

I then subset the data to include only the years I want for the Massachusetts replication, only counties, and add in a variable for an age group (simply the addition of two age groups).

```{r}
pop06=subset(pop99,pop99$year>=2001 & pop99$year<=2006)
pop06=subset(pop06,pop06$type==3)
pop06$a2034=pop06$a2024+pop06$a2534

mort06=subset(mort99,mort99$year>=2001 & mort99$year<=2006)
mort06=subset(mort06,mort06$age>=09 & mort06$age<=13)

```

The following section was added to provide a mortality rate for all years, to help with the analysis of the comparison post-treatment.

```{r}
popall=subset(pop99,pop99$type==3)
popall$a2034=popall$a2024+popall$a2534
mortall=subset(mort99, mort99$age>=09 & mort99$age<=13)

```

As that was the last use of `pop99` in this script:
```{r}
rm(pop99,mort99)
save(popall,mortall,pop06,mort06,arf, file=paste0(datadir, "created/mass_prep.RData"))
```
