---
title: "Massachusetts Propensity Score Model"
author: "Julie Ghekas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

I am documenting the decisions, assumptions, and confidence in those decisions in replicating the analysis for Massachusetts presented in the Sommers et al paper.


Read in subsets of `pop` and `mort` data files (created by `mass_prep.Rmd`).
```{r}
stopifnot(file.exists(prepped_data_file <- paste0(datadir, "created/mass_prep.RData")))
load(prepped_data_file)
library(reshape2)
library(car)
library(withr)
with_libpaths("lib/", library(RItools), "prefix")

```


I then begin aggregating the data to accomplish the variables that I wish to have (in replication of the propensity score model discussed in Appendix Table 2).  

```{r}
agemat=aggregate(cbind(a2034,a3544,a4554,a5564)~county+state,sum,data=pop06)
ages=cbind(agemat[,1:2],100*t(apply(agemat[,3:6],1,prop.table)))

racesexmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~racesex+county+state,sum,data=pop06)
racesmat=racesexmat
racesmat$race=recode(racesexmat$racesex,"1:2=1; 3:4=2; 5:8=3")
racesmat$race[racesmat$race==1]="white"
racesmat$race[racesmat$race==2]="black"
racesmat$race[racesmat$race==3]="otherrace"
races=aggregate(count~race+county+state,sum,data=racesmat)
races=dcast(data=races,state+county~race)
race=cbind(races[,1:2],100*t(apply(races[,3:5],1,prop.table)))

sexmat=racesexmat
sexmat$sex=recode(racesexmat$racesex, "1=1; 2=2; 3=1; 4=2; 5=1; 6=2; 7=1; 8=2")
sexmat$sex[sexmat$sex==1]="male"
sexmat$sex[sexmat$sex==2]="female"
sexes=aggregate(count~sex+county+state,sum,data=sexmat) 
sexes=dcast(data=sexes,state+county~sex)
sex=cbind(sexes[,1:2],100*t(apply(sexes[,3:4],1,prop.table)))

hispmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~hisp+county+state,sum,data=pop06)
hispmat$hisp=recode(hispmat$hisp, '2=1; 1=2; 9=2')
hispmat$hisp[hispmat$hisp==1]="hisp"
hispmat$hisp[hispmat$hisp==2]='nothisp'
hisps=aggregate(count~hisp+county+state,sum,data=hispmat) 
hisps=dcast(data=hisps,state+county~hisp)
hisp=cbind(hisps[,1:2],100*t(apply(hisps[,3:4],1,prop.table)))
```

Assumptions that I have made here: I assumed that the variables for age, race, sex, etc. were aggregate over the 7 years.  That is, for every individual who did not die or age out of 64 in 2000-2005 or 2006 (depending on how exactly the population count was made), that individual is represented in the data 7 times.  (Other options include looking only at this representation for 2006 or 2000).  Looking at just 2006 could make sense, as it would be the year that was most recent to the insurance being fit.

These counts are all for individuals who are between 20 and 64 years of age, as the paper comments that the population of interest is adults in that age range.


```{r}
deaths=aggregate(deaths~county+state+year,sum,data=mort06)
pop06$pops=pop06$a2024+pop06$a2534+pop06$a3544+pop06$a4554+pop06$a5564
pops=aggregate(pops~county+state+year,sum,data=pop06)
total=merge(pops,deaths,by=c('year','state','county'))
total$mrate=total$deaths/total$pops*100000
totals=dcast(data=total,state+county~year,value.var='mrate')
colnames(totals)=c('state','county','mr01','mr02','mr03','mr04','mr05','mr06')
```


Here, I am going through and implementing the (modified) propensity
score model, using only the variables that I have available to me
directly from the CDC data, without using any other sources.  This
relies on the removal of poverty rate, median household income,
unemployment rate, and percent uninsured from the model.
**BH**: Is this comment out of sync with code?  The next PS model
below appears to have the named variables.

```{r}
model2data=merge(ages,race,by=c('state','county'))
model2data=merge(model2data,sex,by=c('state','county'))
model2data=merge(model2data,hisp,by=c('state','county'))
model2data=merge(model2data,totals,by=c('state','county'))
model2data$treat=0
model2data$treat[model2data$state==25]=1
model2data=merge(model2data,arf,by=c('state','county'))
```

There are about 50 NAs in the model2data dataframe, all arising from
`totals`:
```{r}
sum(!complete.cases(model2data))
sum(is.na(model2data))
all.equal(sum(is.na(model2data)), sum(is.na(totals)))
```
The way that the mortalities are recorded, to the best of my understanding, is as a count.  The mortality entry only occurs if at least one individual died of that cause within a given age group, racial characteristic, and gender.  I believe that if a death didn't occur within a given demographic, this would result in a count of NA within the aggregated data.  For some small counties, there may be year(s) without deaths.  So, for a year without deaths, the rate would not be calculable.  Based on this belief, I will record the mortality rate for these counties as 0, resulting in a complete dataframe.  

```{r}
model2data[is.na(model2data)]=0
```

Here is the propensity model:
```{r}
model2=glm(treat~a2034+a3544+a4554+male+white+black+hisp+
               mr01+mr02+mr03+mr04+mr05+mr06+
                   unemp06+pov06+ins06+medinc06,
    data=model2data,family=binomial(link='logit'))
stopifnot(is.null(model2$na.action))
summary(model2)
optmatch:::boxplot.glm(model2)
```

This model predicts that for some counties, there is a probability of 0 that the county is in Massachusetts (the treatment group).  Below, I will explore the fitted values from this model.  The coefficients for this model do not appear to be consistent with those presented in Sommers' paper.  

```{r}
hist(fitted(model2),breaks=60)
summary(fitted(model2))
table(model2data$treat,
      cut(fitted(model2), quantile(fitted(model2),
                                   probs=c(0,.5,.75,1)))
      )
```

Based on the above analysis, it appears as if one of the treated
counties is between the 50th and 75th percentile, with a fitted value
less than the mean, and another one is less than the median.
(The mean is larger than Q3, which makes sense with the skewed right shape).  

```{r}
fit.q3=summary(fitted(model2))[5]
cont.count=which(fitted(model2)>fit.q3)
length(unique(model2data$state[cont.count]))
```

```{r}
mass.study = subset(model2data, state==25 | fitted(model2)>fit.q3)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+
             white+male+hisp+mr01+mr02+mr03+mr04+mr05+mr06,
         data=mass.study,report=c('all'))
```

## First look at outcomes

This section is added below to be able to compare the outcomes amongst the groups

```{r}
deathall=aggregate(deaths~county+state+year, sum, data=mortall)
popall$pops=popall$a2024+popall$a2534+popall$a3544+popall$a4554+popall$a5564
popall=aggregate(pops~county+state+year,sum,data=popall)
totalall<-merge(popall,deathall,by=c('year','state','county'))
totalall$mrate=totalall$deaths/totalall$pops*100000
totalsall=dcast(data=totalall,state+county~year,value.var='mrate')
colnames(totalsall)=c('state','county','mr99','mr00','mr01','mr02','mr03','mr04','mr05','mr06','mr07','mr08','mr09','mr10','mr11','mr12','mr13')
sapply(totalsall[1:2], function(x) any(is.na(x)) )
totalsall[is.na(totalsall)]=0
```

Assumptions that I have made here: I assumed that this was general mortality.  I did not calculate the deaths amenable to health care rate, just the all told mortality rate.  I also assumed that the mortality rate only had individuals 20-64 considered.


This chunk aims to compare outcomes between the groups.

```{r}
mass.control.mort<-totalsall[cont.count,]
mass.control.mort<-subset(mass.control.mort,mass.control.mort$state!=25)
mass.treat.mort<-subset(totalsall,totalsall$state==25)
mass.mort.study<-rbind(mass.treat.mort,mass.control.mort)
mass.mort.study$treat=0
mass.mort.study$treat[mass.mort.study$state==25]=1
xBalance(treat~mr07+mr08+mr09+mr10+mr11+mr12+mr13,data=mass.mort.study,report=c('all'))
```

...but note that there's a premise here that `model2data` is aligned
with `totalsall`, and that premise turns out to be mistaken:
```{r}
nrow(model2data)
nrow(totalsall)
```

So the results should be taken with a grain of salt.
