---
title: "Massachusetts Outcome Analysis Exploration"
author: "Julie Ghekas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

Outcome analysis is performed here for Massachusetts.  The matching scheme that we 
selected used an unweighted glm to generate the propensity scores, full matching 
retaining 1/4 of the potential control counties, and a population-weighted balance 
assessment.  The groups formed from this procedure will function as the two groups to 
be compared to assess if the outcomes are different.  The full pool of potential 
comparisons, i.e. all counties not in Massachusetts, may be used at various points in 
time, but these are not the primary comparisons of interest.

```{r}
stopifnot(file.exists(prepped_data_file <- paste0(datadir, "created/mass_prep.RData")))
load(prepped_data_file)

library(reshape2)
library(car)
library(optmatch)
library(brglm)
library(MASS)
library(geepack)

stopifnot(file.exists("./lib"))
if ( file.exists(paste0("./lib/", R.version[['platform']], "-library") ) )
{
  with_libpaths(with(R.version, paste0("./lib/", platform, "-library/", major, ".", substr(minor, 1,1))),
		library(RItools)
		)
} else with_libpaths("lib/", library(RItools))
```

```{r}
agemat=aggregate(cbind(a2034,a3544,a4554,a5564)~county+state,sum,data=pop06)
ages=cbind(agemat[,1:2],100*t(apply(agemat[,3:6],1,prop.table)))

racesexmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~racesex+county+state,sum,data=pop06)
racesmat=racesexmat
racesmat$race=recode(racesexmat$racesex,"1:2=1; 3:4=2; 5:8=3")
racesmat$race[racesmat$race==1]="white"
racesmat$race[racesmat$race==2]="black"
racesmat$race[racesmat$race==3]="otherrace"
races=aggregate(count~race+county+state,sum,data=racesmat)
races=dcast(data=races,state+county~race, value.var='count')
race=cbind(races[,1:2],100*t(apply(races[,3:5],1,prop.table)))

sexmat=racesexmat
sexmat$sex=recode(racesexmat$racesex, "1=1; 2=2; 3=1; 4=2; 5=1; 6=2; 7=1; 8=2")
sexmat$sex[sexmat$sex==1]="male"
sexmat$sex[sexmat$sex==2]="female"
sexes=aggregate(count~sex+county+state,sum,data=sexmat) 
sexes=dcast(data=sexes,state+county~sex, value.var = 'count')
sex=cbind(sexes[,1:2],100*t(apply(sexes[,3:4],1,prop.table)))

hispmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~hisp+county+state,sum,data=pop06)
hispmat$hisp=recode(hispmat$hisp, '2=1; 1=2; 9=2')
hispmat$hisp[hispmat$hisp==1]="hisp"
hispmat$hisp[hispmat$hisp==2]='nothisp'
hisps=aggregate(count~hisp+county+state,sum,data=hispmat) 
hisps=dcast(data=hisps,state+county~hisp, value.var='count')
hisp=cbind(hisps[,1:2],100*t(apply(hisps[,3:4],1,prop.table)))

deaths=aggregate(deaths~county+state+year,sum,data=mort06)

pop06$pops=pop06$a2024+pop06$a2534+pop06$a3544+pop06$a4554+pop06$a5564
pops=aggregate(pops~county+state+year,sum,data=pop06)
popscheck = dcast(data=pops, state+county~year, value.var='pops')
popscheck[which(is.na(popscheck$'2006')),]
rbind(popscheck[which(is.na(popscheck$'2001')),], 
      popscheck[which(is.na(popscheck$'2002')),], 
      popscheck[which(is.na(popscheck$'2003')),], 
      popscheck[which(is.na(popscheck$'2004')),], 
      popscheck[which(is.na(popscheck$'2005')),], 
      popscheck[which(is.na(popscheck$'2006')),])
total=merge(pops,deaths,by=c('year','state','county'), all=T)
total$mrate=total$deaths/total$pops*100000
total[which(is.na(total$mrate)),]
all.equal(which(is.na(total$deaths)),which(is.na(total$mrate)))
totals=dcast(data=total,state+county~year,value.var='mrate')
colnames(totals)=c('state','county','mr01','mr02','mr03','mr04','mr05','mr06')

totals[is.na(totals)] = 0
pops=dcast(data=total,state+county~year,value.var='pops')
pops$pop06 <- apply(pops[ncol(pops):3L], 1, function(x) x[!is.na(x)][1])
all.equal(pops[!is.na(pops$'2006'),'2006'],pops[!is.na(pops$'2006'),'pop06'])
pops <- subset(pops, select=c(state,county,pop06))
model2data=merge(ages,race,by=c('state','county'))
model2data=merge(model2data,sex,by=c('state','county'))
model2data=merge(model2data,hisp,by=c('state','county'))
model2data=merge(model2data,totals,by=c('state','county'))
model2data=merge(model2data,pops,by=c('state','county'))
model2data$treat=0
model2data$treat[model2data$state==25]=1
model2data=merge(model2data,arfall,by=c('state','county'))

```

```{r}
model2=glm(treat~a2034+a3544+a4554+male+white+black+hisp+mr01+mr02+mr03+mr04+mr05+mr06+
             pov06+medinc06+unemp06+ins06,data=model2data,family=binomial(link='logit'))
fm2<-fullmatch(model2,data=model2data,omit.fraction=3/4)
model2fit = cbind(model2data, fm2)
model2fit = model2fit[,c(1,2,31)]

popyr = subset(popall, popall$year>=2001)
popyr$pops = popyr$a2034 + popyr$a3544 + popyr$a4554 + popyr$a5564
popsyr = aggregate(pops~county+state+year, sum, data=popyr)

ageyr = aggregate(cbind(a2034, a3544, a4554, a5564)~county+state+year, sum, data=popyr)
ageyr = cbind(ageyr[,1:3],100*t(apply(ageyr[,4:7],1,prop.table)))

rsyr = aggregate(cbind(count=a2034+a3544+a4554+a5564)~racesex+county+state+year,sum,
                 data=popyr)
ryr = rsyr
ryr$race = recode(ryr$racesex, "1:2=1; 3:4=2; 5:8=3")
ryr$race[ryr$race==1] = 'white'
ryr$race[ryr$race==2] = 'black'
ryr$race[ryr$race==3] = 'otherrace'
racesyr = aggregate(count~race+county+state+year, sum, data=ryr)
racesyr = dcast(data=racesyr, state+county+year~race)
raceyr = cbind(racesyr[,1:3],100*t(apply(racesyr[,4:6],1,prop.table)))

syr = rsyr
syr$sex = recode(syr$racesex, "1=1; 2=2; 3=1; 4=2; 5=1; 6=2; 7=1; 8=2")
syr$sex[syr$sex==1]='male'
syr$sex[syr$sex==2]='female'
sexesyr = aggregate(count~sex+county+state+year, sum, data=syr)
sexesyr = dcast(data=sexesyr, state+county+year~sex)
sexyr = cbind(sexesyr[,1:3], 100*t(apply(sexesyr[,4:5],1,prop.table)))

hispyr = aggregate(cbind(count=a2034+a3544+a4554+a5564)~hisp+county+state+year, sum, 
                   data=popyr)
hispyr$hisp = recode(hispyr$hisp, '2=1;1=2;9=2')
hispyr$hisp[hispyr$hisp==1] = 'hisp'
hispyr$hisp[hispyr$hisp==2] = 'nothisp'
hispsyr = aggregate(count~hisp+county+state+year, sum, data=hispyr)
hispsyr = dcast(data=hispsyr, state+county+year~hisp)
hispyr = cbind(hispsyr[,1:3],100*t(apply(hispsyr[,4:5],1,prop.table)))

mortyr = subset(mortall, mortall$year>=2001)
deathsyr = aggregate(deaths~county+state+year, sum, data=mortyr)

yeardata = merge(deathsyr, popsyr, by=c('state','county','year'))
yeardata = merge(yeardata, ageyr, by=c('state','county','year'))
yeardata = merge(yeardata, raceyr, by=c('state','county','year'))
yeardata = merge(yeardata, sexyr, by=c('state','county','year'))
yeardata = merge(yeardata, hispyr, by=c('state','county','year'))
yeardata$treat = 0
yeardata$treat[yeardata$state==25 & yeardata$year>=2006] = 1
yeardata$idvar = paste(yeardata$state, yeardata$county, sep=',')


year.group = merge(yeardata, model2fit, by=c('state','county'), all=T)
year.group = year.group[order(year.group$state, year.group$county, year.group$year),]
year.group = subset(year.group, !is.na(year.group$fm2))
year.group$idvar = as.factor(year.group$idvar)

massgee = geese(deaths/pops~a2034+a3544+a4554+black+otherrace+female+hisp+treat+
                  strata(fm2), id=idvar, family=gaussian, data=year.group, waves=year)
summary(massgee)
```

Below are visualizations for Massachusetts, the control group, and all counties outside
of Massachusetts for a handful of variables, including insurance and mortality.

```{r}
#I need to perform weighted means for the control group

total=merge(popsyr,deathsyr,by=c('year','state','county'), all=T)
total$mrate=total$deaths/total$pops*100000
totals=dcast(data=total,state+county~year,value.var='mrate')
colnames(totals)=c('state','county','mr01','mr02','mr03','mr04','mr05','mr06','mr07',
                   'mr08','mr09','mr10','mr11','mr12','mr13')
totals[is.na(totals)]=0
totals = subset(totals, select = -c(mr01:mr06))


deathdata6 = merge(model2data, arfall, by=c('state','county'))
deathdata6 = merge(deathdata6, totals, by=c('state','county'))
colnames(deathdata6)
mass.group = merge(deathdata6, model2fit, by=c('state','county'))
mass.group = subset(mass.group, !is.na(fm2))
all.mort = apply(deathdata6[,c(14:19,69:75)],2,mean)
cont.mort = apply(subset(mass.group,mass.group$treat==0)[,c(14:19,69:75)],2,mean)
mass.mort = apply(subset(deathdata6,deathdata6$treat==1)[,c(14:19,69:75)],2,mean)
allm.mort = apply(deathdata6[,c(14:19,69:75)],2,median)
contm.mort = apply(subset(mass.group,mass.group$treat==0)[,c(14:19,69:75)],2,median)
massm.mort = apply(subset(deathdata6,deathdata6$treat==1)[,c(14:19,69:75)],2,median)

plot(c(1:13),mass.mort,ylim=c(280,440),type='l',col='red')
lines(c(1:13),cont.mort,col='purple')
lines(c(1:13),all.mort,col='blue')
lines(c(1:13),massm.mort,type='l',col='red',lty=2)
lines(c(1:13),allm.mort,type='l',col='blue',lty=2)
lines(c(1:13),contm.mort,type='l',col='purple',lty=2)
all.ins = apply(deathdata6[,c(46,27,64,63,62,26)],2,mean)
cont.ins = apply(subset(mass.group,mass.group$treat==0)[,c(46,27,64,63,62,26)],2,mean)
mass.ins = apply(subset(deathdata6,deathdata6$treat==1)[,c(46,27,64,63,62,26)],2,mean)
allm.ins = apply(deathdata6[,c(46,27,64,63,62,26)],2,median)
contm.ins = apply(subset(mass.group, mass.group$treat==0)[,c(46,27,64,63,62,26)],
                  2, median)
massm.ins = apply(subset(deathdata6, deathdata6$treat==1)[,c(46,27,64,63,62,26)], 
                  2, median)
plot(c(5:10),mass.ins,ylim=c(5,23),type='l',col='red')
lines(c(5:10),cont.ins,col='purple')
lines(c(5:10),all.ins,col='blue')
lines(c(5:10),massm.ins,col='red',lty=2)
lines(c(5:10),contm.ins,col='purple',lty=2)
lines(c(5:10),allm.ins,col='blue',lty=2)

```


The ARF data appear to have no missing variables.  In fact, 0s have been imputed for missing variables.  In most cases, it appears as if data are missing for the county as a whole, not for a specific year.  This will need to be addressed.

As a double check of the strata command within the geese function:
```{r}
ab = matrix(rnorm(500),nrow=100,ncol=5)
ab = as.data.frame(ab)
idvar = c(rep(1, 10), rep(2, 10), rep(3, 10), rep(4, 10), rep(5, 10), rep(6, 10), 
          rep(7, 10), rep(8, 10), rep(9, 10), rep(10, 10))
ce = c(rep('1.1', 5), rep('1.2', 6), rep('1.3', 7), rep('1.4', 8), rep('1.5', 9), 
       rep('1.6', 10), rep('1.7', 11), rep('1.8', 12), rep('1.9', 13), rep('1.10', 14), 
       rep('1.11', 5))
ce = as.factor(ce)
fg = matrix(0,nrow=100,ncol=11)
fg[c(1:5),1] = 1
fg[c(6:11),2] = 1
fg[c(12:18),3] = 1
fg[c(19:26),4] = 1
fg[c(27:35),5] = 1
fg[c(36:45),6] = 1
fg[c(46:56),7] = 1
fg[c(57:68),8] = 1
fg[c(69:81),9] = 1
fg[c(82:95),10] = 1
fg[c(96:100),11] = 1
colnames(ab) = c('a1','b2','c3','d4','e5')
colnames(fg) = c('s1','s2','s3','s4','s5','s6','s7','s8','s9','s10','s11')
ab1 = cbind(ab,ce,idvar)
ab2 = cbind(ab,fg,idvar)


m11 = geese(a1~b2+c3+d4+e5+strata(ce),data=ab1, id=idvar)
m12 = geese(a1~b2+c3+d4+e5+s2+s3+s4+s5+s6+s7+s8+s9+s10+s11, id=idvar, data=ab2)
summary(m11)
summary(m12)
```