---
title: "Tennessee Code"
author: "Julie Ghekas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This file provides the matching schemes and their balance assessments.  The first chunk 
provides the appropriate setup and configuration of the computer.  This chunk relies on 
mass_prep having been run to load in the appropriate versions of the data.  The second 
chunk calculates and configures the data into the appropriate format.  The calculations 
performed here are the same as those performed for the Kentucky analysis until it comes 
to including counties to serve as the control counties.  The third chunk provides the 
matching schemes that were considered with different adjustments.  The adjustments 
examined are: 4 models (brglm vs glm, population weighted vs unweighted), 3 
matching/control group schemes (top quartile propensity score, full matching with 1/4 
of the potential control group retained, and pair matching), and 2 balance assessments 
(population weighted vs unweighted), for a total of 24 different balance assessments 
within this document. 

```{r}
stopifnot(file.exists(prepped_data_file <- paste0(datadir, "created/mass_prep.RData")))
load(prepped_data_file)

stopifnot(file.exists("./lib"))
if ( file.exists(paste0("./lib/", R.version[['platform']], "-library") ) )
{
  with_libpaths(with(R.version, paste0("./lib/", platform, "-library/", major, ".", substr(minor, 1,1))),
		library(RItools)
		)
} else with_libpaths("lib/", library(RItools))

library(reshape2)
library(car)
library(optmatch)
```

```{r}
pop13=subset(popall, popall$year>=2005)
agemat=aggregate(cbind(a2034,a3544,a4554,a5564)~county+state,sum,data=pop13)
ages=cbind(agemat[,1:2],100*t(apply(agemat[,3:6],1,prop.table)))

racesexmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~racesex+county+state,sum,data=pop13)
racesmat=racesexmat
racesmat$race=recode(racesexmat$racesex,"1:2=1; 3:4=2; 5:8=3")
racesmat$race[racesmat$race==1]="white"
racesmat$race[racesmat$race==2]="black"
racesmat$race[racesmat$race==3]="otherrace"
races=aggregate(count~race+county+state,sum,data=racesmat)
races=dcast(data=races,state+county~race)
race=cbind(races[,1:2],100*t(apply(races[,3:5],1,prop.table)))

sexmat=racesexmat
sexmat$sex=recode(racesexmat$racesex, "1=1; 2=2; 3=1; 4=2; 5=1; 6=2; 7=1; 8=2")
sexmat$sex[sexmat$sex==1]="male"
sexmat$sex[sexmat$sex==2]='female'
sexes=aggregate(count~sex+county+state,sum,data=sexmat) 
sexes=dcast(data=sexes,state+county~sex)
sex=cbind(sexes[,1:2],100*t(apply(sexes[,3:4],1,prop.table)))

hispmat=aggregate(cbind(count=a2034+a3544+a4554+a5564)~hisp+county+state,sum,data=pop13)
hispmat$hisp=recode(hispmat$hisp, "2=1; 1=2; 9=2")
hispmat$hisp[hispmat$hisp==1]="hisp"
hispmat$hisp[hispmat$hisp==2]='nothisp'
hisps=aggregate(count~hisp+county+state,sum,data=hispmat) 
hisps=dcast(data=hisps,state+county~hisp)
hisp=cbind(hisps[,1:2],100*t(apply(hisps[,3:4],1,prop.table)))
mort13=subset(mortall,mortall$year>=2005)
deaths=aggregate(deaths~county+state+year,sum,data=mort13)
pop13$pops=pop13$a2024+pop13$a2534+pop13$a3544+pop13$a4554+pop13$a5564
pops=aggregate(pops~county+state+year,sum,data=pop13)
total=merge(pops,deaths,by=c('year','state','county'))
total$mrate=total$deaths/total$pops*100000
totals=dcast(data=total,state+county~year,value.var='mrate')
colnames(totals)=c('state','county','mr05','mr06','mr07','mr08','mr09','mr10','mr11','mr12','mr13')
totals[is.na(totals)]=0
model3data=merge(ages,race,by=c('state','county'))
model3data=merge(model3data,sex,by=c('state','county'))
model3data=merge(model3data,hisp,by=c('state','county'))
model3data=merge(model3data,totals,by=c('state','county'))
model3data=merge(model3data,arf,by=c('state','county'))
model4data=model3data
model4data$treat=0
model4data$treat[model4data$state==47]=1
states.tncontrol=c('4','5','6','8','9','10','17','18','19','21','24','25','26','27','30','32','33','34','35','36','38','39','41','42','44','47','50','53','54')
model4data=model4data[model4data$state %in% states.tncontrol,]
```

The control counties above were located in states that had adopted changes to their 
Medicaid program as a result of the ACA.  The source and citation for this list can be 
found in the ky_code.Rmd file.

The various matching schemes and control groups are included below.

```{r}
model4=glm(treat~a2034+a3544+a4554+male+white+black+hisp+mr05+mr06+mr07+mr08+mr09+mr10+
             pov10+medinc10+unemp10+ins10,data=model4data,family=binomial(link='logit'))
model4a=glm(treat~a2034+a3544+a4554+male+white+black+hisp+mr05+mr06+mr07+mr08+mr09+mr10+
              pov10+medinc10+unemp10+ins10,data=model4data,
            family=binomial(link='logit'),weights=pops)
model4b=brglm(treat~a2034+a3544+a4554+male+white+black+hisp+mr05+mr06+mr07+mr08+mr09
              +mr10+
                pov10+medinc10+unemp10+ins10,data=model4data,
              family=binomial(link='logit'))
model4c=brglm(treat~a2034+a3544+a4554+male+white+black+hisp+mr05+mr06+mr07+mr08+
                mr09+mr10+
                pov10+medinc10+unemp10+ins10,data=model4data,
              family=binomial(link='logit'),weights=pops)
fit=fitted(model4)
fita=fitted(model4a)
fitb=fitted(model4b)
fitc=fitted(model4c)
model4data=cbind(model4data,fit,fita,fitb,fitc)
optmatch:::boxplot.glm(model4)
optmatch:::boxplot.glm(model4a)
optmatch:::boxplot.glm(model4b)
optmatch:::boxplot.glm(model4c)

#model 4: glm unweighted
ten.treat=subset(model4data,model4data$treat==1)
ten.control=subset(model4data,model4data$treat==0)
top4=summary(model4data$fit)[5]
ten.control=subset(ten.control,ten.control$treat!=1)
ten.control=subset(ten.control, ten.control$fit>top4)
ten.study=rbind(ten.treat,ten.control)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, data=ten.study,report=c('all'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, element.weights=pops, 
         data=ten.study,report=c('all'))

fm4<-fullmatch(model4,data=model4data,omit.fraction=3/4)
summary(fm4)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)

pm_ten<-pairmatch(model4,data=model4data)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_ten),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_ten),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)
stratumStructure(pm_ten)

#####model4a: glm weighted
ten.treat=subset(model4data,model4data$treat==1)
ten.control=subset(model4data,model4data$treat==0)
top4=summary(model4data$fita)[5]
ten.control=subset(ten.control,ten.control$treat!=1)
ten.control=subset(ten.control, ten.control$fita>top4)
ten.study=rbind(ten.treat,ten.control)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, data=ten.study,report=c('all'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, element.weights=pops, 
         data=ten.study,report=c('all'))

fm4a<-fullmatch(model4a,data=model4data,omit.fraction=3/4)
summary(fm4a)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4a),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4a),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)

pm_tena<-pairmatch(model4a,data=model4data)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tena),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tena),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)
stratumStructure(pm_tena)

###model4b: brglm unweighted
ten.treat=subset(model4data,model4data$treat==1)
ten.control=subset(model4data,model4data$treat==0)
top4=summary(model4data$fitb)[5]
ten.control=subset(ten.control,ten.control$treat!=1)
ten.control=subset(ten.control, ten.control$fitb>top4)
ten.study=rbind(ten.treat,ten.control)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, data=ten.study,report=c('all'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, element.weights=pops, 
         data=ten.study,report=c('all'))

fm4b<-fullmatch(model4b,data=model4data,omit.fraction=3/4)
summary(fm4b)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4b),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4b),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)

pm_tenb<-fullmatch(model4b,data=model4data,min.controls=1,max.controls=1)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tenb),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tenb),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)
stratumStructure(pm_tenb)

###model4c: brglm weighted
ten.treat=subset(model4data,model4data$treat==1)
ten.control=subset(model4data,model4data$treat==0)
top4=summary(model4data$fitc)[5]
ten.control=subset(ten.control,ten.control$treat!=1)
ten.control=subset(ten.control, ten.control$fitc>top4)
ten.study=rbind(ten.treat,ten.control)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, data=ten.study,report=c('all'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10, element.weights=pops, 
         data=ten.study,report=c('all'))

fm4c<-fullmatch(model4c,data=model4data,omit.fraction=3/4)
summary(fm4c)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4c),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(fm4c),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)

pm_tenc<-fullmatch(model4c,data=model4data,min.controls=1,max.controls=1)
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tenc),data=model4data,
         report=c('z','chisquare.test'))
xBalance(treat~a2034+a3544+a4554+a5564+black+otherrace+white+male+hisp+mr05+mr06+mr07+
           mr08+mr09+mr10+pov10+medinc10+unemp10+ins10+strata(pm_tenc),data=model4data,
         report=c('z','chisquare.test'),element.weights=pops)
stratumStructure(pm_tenc)
```